Human Factors Influencing Visual Statistical Inference

### 2013-06-08 sociology_paper.tex
updated overview of the data

\subsection{Data collection methods}  The human observers who evaluate the experimental lineups are recruited through Amazon Mechanical Turk \cite{turk} or MTurk Web site.  It is an online work place where people from around the world can perform some tasks and get paid. Usually tasks are very simple and no specialized training is required. Being a human is the main requirement. Tasks are designed for anyone to do but some tasks may require that workers satisfy some skill level depending on the recruiters need. The tasks are designed such that it does not take much time to complete. Humans are still better than computers in performing these types of tasks. The the amount of money paid for each task is very small as well.

@@

Because of its convenience it is getting popular for scientific research study. In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience. \cite{majumder:2013} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference. They have done numerous pilot studies in lab before doing actual MTurk study and found similar results. \cite{mason:2012} explains various features of MTurk and describes how it can be used as part of human behavioral study.

We design and develop a web site which enables to display the lineups to the observers as per experimental design. The MTurk workers are redirect to the web site and the data are collected and automatically stored into a local server. Demographic informations such as age group, gender and education levels were collected. The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback is received. The location of the observer is determined by the ip address of the observer.

@@

\subsection{Overview of the Data} Figure \ref{fig:turker_location} shows the location from where we received data. Notice that we received data from around the world and we observed diversity of location in the collected data. We notice diversity in not only location of the observers but also their gender, age groups and education levels. Table \ref{tbl:demographics} shows that number of male and female participants are almost same.

@@

The large number of participants are from age group 18 to 35. Interestingly we see many participants from older age groups as well. Specially for united states almost all the age groups show uniform participations after age 30 as in Figure \ref{fig:demographic_info}. Notice that fewer people participated from India beyond age 40 compared to united states. Even thoug total participations from india is larger compared to United States, they are mostly young people who are capable of using internet and computer.

\subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken for each lineup are computed and their distributions are shown in Figure \ref{fig:demographic_effect} for different demographic factor variables.

@@

@@

More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of statistical graphics. The numerous pilot studies done with the more trained participants suggest that power of visual statical inference may be higher with skilled observer. Thus the variations observed with individual performances could be attributed to individual abilities.

>>>>

\subsection{Data collection methods}  Human observers were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \cite{turk} or MTurk Web site.  It is an online work place where people from around the world can perform some tasks and get paid. Usually tasks are very simple and no specialized training is required. Being a human is the main requirement. Tasks are designed for anyone to do but some tasks may require that workers satisfy some skill level depending on the recruiters need. The tasks are designed such that it does not take much time to complete. Humans are still better than computers in performing these types of tasks. The the amount of money paid for each task is very small as well.

@@

Because of its convenience it is getting popular for scientific research study as well. In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience. \cite{majumder:2013} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference. They have done numerous pilot studies in lab before doing actual MTurk study and found similar results. \cite{mason:2012} explains various features of MTurk and describes how it can be used as part of human behavioral study.

We design and develop a web site which enables to display the lineups to the observers as per experimental design. The MTurk workers are redirect to the web site and the data are collected, stored automatically into a local database server. Demographic informations such as age group, gender and education levels were collected. The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received. The location of the observer is determined by the ip address of the observer.

@@

\subsection{Overview of the Data} Figure \ref{fig:turker_location} displays the location around the world from where a total of 2321 participants provided feedback data on the lineups in all the ten experimental studies. Even though most of the participants were from the United States and India, we have participants from 76 different countries. This provides a diverse pool of participants in terms of geographical locations. The diversity in not only in geographical locations of the participants but also in their gender, age groups and education levels as we see in Table \ref{tbl:demographics}.  It is interesting that the number of male and female participants are almost same even though there are lot of people from developing countries.

Besides United States and India, countries such as Canada, Romania, United Kingdom and Macedonia have more than 10 participants each. The rest of the 70 countries have less than 10 participants.

@@

The largest number of participants are from age group 18 to 25 which is the youngest age group in the study. Interestingly we see many participants from older age groups as well. Specially for united states almost all the age groups show uniform participations after age 30 as in Figure \ref{fig:demographic_info}. Notice that fewer people participated from India beyond age 40 compared to united states. Even thoug total participations from india is larger compared to United States, they are mostly young people who are capable of using internet and computer.

\subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken for each lineup are computed and their distributions are shown as boxplots in Figure \ref{fig:demographic_effect} for different demographic factor levels.

@@

Observer demographics don't seem to have significant influence on performance in time taken to evaluate a lineup. Some variation is observed in terms of median proportion correct, but the mean proportion correct over different demographic factors levels seem to be similar.

@@

More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of knowledge about statistical graphics. The subjects participating this study may not necessarily knowledgable about statistical graphics. The numerous pilot studies done with the more trained participants suggest that power of visual statical inference may be higher with knowledgeable observer. Thus the variations observed with individual performances could be attributed to individual abilities in terms of their training on statistical graphics.
